Density dependence

### 2020-09-24 R/density dependence.Rmd
messing with gams

But how bad is this convergence failure?

Is the max|grad| calculation correct?

```{r}
derivs1 <- mnorm@optinfo$derivs
sc_grad1 <- with(derivs1, solve(Hessian, gradient))
max(abs(sc_grad1))
```
No, it's actually a lot smaller than `tol`, so nothing to worry about

@@

```{r}
m_growth2 <- glmer(shts_next ~ log_shts + habitat + shts_0_5 + shts_5_10 + (1|year_post),
               family = gaussian(link = "log"),
               data = ha_sub)
AICtab(m_growth, m_growth2)
```

@@

```{r}
m_growth3 <- glmer(shts_next ~ log_shts + habitat + shts_0_5 + shts_5_10 + (1|ha_id_number),
               family = gaussian(link = "log"),
               data = ha_sub)
AICtab(m_growth, m_growth2, m_growth3)
```

@@

```{r}
mgrowth_sub <- glmer(shts_next ~ log_shts + habitat + shts_0_5 + shts_5_10 + (1|ha_id_number) + (1|year_post),
                family = gaussian(link = "log"),
                data = ha_sub_sub)
Anova(mgrowth_sub)
```

@@

```{r}
m_surv2 <- glmer(surv ~ log_shts + habitat + shts_0_5 + shts_5_10 + (1|year), 
                family = binomial, data = ha_sub)
AICtab(m_surv, m_surv2)
```

@@

@@

Marginal effects on survival.

@@

```{r}
m_surv2_sub <- glmer(surv ~ log_shts + habitat + shts_0_5 + shts_5_10 + (1|year), 
                family = binomial, data = ha_sub_sub)
Anova(m_surv2_sub)
```

No longer significant

@@

```{r}
m_flwr <- glmer(flwr ~ log_shts + habitat + shts_0_5 + shts_5_10 + (1|ha_id_number) + (1|year),
                family = binomial, data = ha_sub)
```

@@

```{r}
m_flwr2 <- glmer(flwr ~ log_shts + habitat + shts_0_5 + shts_5_10 + (1|year),
                 family = binomial, data = ha_sub)
AICtab(m_flwr, m_flwr2)
```
Can't remove plant ID.

```{r}
m_flwr3 <- glmer(flwr ~ log_shts + habitat + shts_0_5 + shts_5_10 + (1|ha_id_number),
                 family = binomial, data = ha_sub)
AICtab(m_flwr, m_flwr2, m_flwr3)
```

@@

```{r eval=FALSE}
Q <-
  ha_sub %>%
  select(density_0_10:density_50_60) %>%
  as.matrix()

L <-
  matrix(seq(5, 55, length.out = 6),
         nrow = nrow(Q),
         ncol = 6,
         byrow = TRUE)
```

@@

# Q_log <- log(Q+1)
g_growth <-
  gam(shts_next ~ s(log_shts) +
        habitat +
        s(Q, L,
          bs = "cb",
          k = c(5, 4),
          xt = list(bs = "cr")),
      family = gaussian(link = "log"),
      data = ha_sub,
      method = "REML")
```

```{r eval=FALSE}
g_surv <-
  gam(surv ~ s(log_shts) +
        habitat +
        s(Q, L,
          bs = "cb",
          k = c(5, 4),
          xt = list(bs = "cr")),
      family = binomial,
      data = ha_sub,
      method = "REML")
```

@@

```{r eval=FALSE}
# x <- crosspred("Q", model = g_surv, cumul = TRUE)
x <- crosspred("Q", model = g_growth, cumul = TRUE)
summary(x)
```

@@

```{r eval=FALSE}
plot.crosspred(x,
               ptype = "contour",
               xlab = "#shoots/100m^2",
               ylab = "distance (m)",
               cumul = FALSE)
```
So this shows negative density dependence on growth that seems to be strongest at about 40-50 m away from the focal plant.

@@

```{r eval=FALSE}
plot.crosspred(x, ptype = "slice", var = 100, xlab = "distance (m)", main = "Effect of 100 shoots/100m^2")
plot.crosspred(x, ptype = "slice", var = 10, xlab = "distance (m)", main = "Effect of 10 shoots/100m^2")
```

### Cumulative

I think this shows the cumulative effect of density summed/integrated over all distances.  So, for example, a plant with about 30 shoots/100m^2 surrounding it out to the maximum distance (60m) would have the largest number of shoots, on average.

```{r eval=FALSE}
plot(x, ptype = "overall", xlab = "density (shoots/100m^2)", cumul = TRUE)
```

There is no built-in function to get cumulative effects of density over all distances.

>>>>

@@

Plant ID:

```{r}
m_growth2 <- update(m_growth, .~. - (1|ha_id_number))
AICtab(m_growth, m_growth2)
```

@@

Year:

```{r}
m_growth3 <- update(m_growth, .~. - (1|year_post))
AICtab(m_growth, m_growth2, m_growth3)
```
Can't remove year

@@

```{r}
mgrowth_sub <- update(m_growth, data = ha_sub_sub)
Anova(mgrowth_sub)
```

@@

Plant ID:

```{r}
m_surv2 <- update(m_surv, .~. - (1|ha_id_number))
AICtab(m_surv, m_surv2)
```

@@

Year:

@@

Positive density dependence for survival.

@@

```{r}
m_surv2_sub <- update(m_surv2, data = ha_sub_sub)
Anova(m_surv2_sub)
fixef(m_surv2_sub)
```

Yes, nearly identical coefs.

@@

```{r}
m_flwr <- glmer(flwr ~ log_shts + habitat + shts_0_5 + shts_5_10 + (1|ha_id_number) + (1|year_post),
                family = binomial, data = ha_sub)
```

@@

Plant ID:

```{r}
m_flwr2 <- update(m_flwr, .~. - (1|ha_id_number))
AICtab(m_flwr, m_flwr2)
```
Can't remove plant ID.

Year:

```{r}
m_flwr3 <- update(m_flwr, .~. - (1|year_post))
AICtab(m_flwr, m_flwr2, m_flwr3)
```

@@

```{r eval=FALSE}
Q <-
  ha_sub %>%
  select(shts_0_0.5:shts_4.5_5) %>%
  as.matrix()

L <-
  matrix(seq(0.25, 4.75, by = 0.5), #use midpoint of r1 and r2
         nrow = nrow(Q),
         ncol = ncol(Q),
         byrow = TRUE)
```

@@

# Q_log <- log(Q+1)
g_growth <-
  gam(shts_next ~ s(log_shts) +
        habitat +
        s(Q, L,
          bs = "cb",
          # k = c(10, 10),
          xt = list(bs = "cr")),
      family = gaussian(link = "log"),
      data = ha_sub,
      method = "REML")
gam.check(g_growth)
anova(g_growth)
```

Marginally significant effect of conspecific density on growth

```{r eval=FALSE}
g_surv <-
  gam(surv ~ s(log_shts, k = 12) +
        habitat +
        s(Q, L,
          bs = "cb",
          # k = c(10, 10),
          xt = list(bs = "cr")),
      family = binomial,
      data = ha_sub,
      method = "REML")
gam.check(g_surv)
anova(g_surv)
```
No significant effect of density

@@

```{r eval=FALSE}
hist(Q)
pred_growth <- crosspred("Q", model = g_growth, cumul = TRUE, cen = mean(Q, na.rm = T), from = 0, to = 8)
summary(pred_growth)
```

@@

```{r}
plot.crosspred(pred_growth,
               ptype = "contour",
               xlab = "#shoots/m^2",
               ylab = "distance (m)",
               cumul = FALSE,
               main = "Relative to mean competition")
```
Non-monotonic density dependence that decays steeply beyond 1m

@@

```{r eval=FALSE}
plot.crosspred(pred_growth, ptype = "slice", var = 0.2, xlab = "distance (m)", main = "0.2 shoots/m^2", ylim = c(0.9,1.1))
plot.crosspred(pred_growth, ptype = "slice", var = 1, xlab = "distance (m)", main = "1 shoots/m^2", ylim = c(0.9,1.1))
plot.crosspred(pred_growth, ptype = "slice", var = 2, xlab = "distance (m)", main = "2 shoots/m^2", ylim = c(0.9,1.1))
plot.crosspred(pred_growth, ptype = "slice", var = 8, xlab = "distance (m)", main = "8 shoots/m^2", ylim = c(0.9,1.1))
```

```{r}
plot.crosspred(pred_growth, ptype = "slice", lag = 0.25, xlab = "density (shoots/m^2)", main = "0.25m", ylim = c(0.9,1.1))
plot.crosspred(pred_growth, ptype = "slice", lag = 1.25, xlab = "density (shoots/m^2)", main = "1.25m", ylim = c(0.9,1.1))
plot.crosspred(pred_growth, ptype = "slice", lag = 3.25, xlab = "density (shoots/m^2)", main = "3.25m", ylim = c(0.9,1.1))
plot.crosspred(pred_growth, ptype = "slice", lag = 4.25, xlab = "density (shoots/m^2)", main = "4.25m", ylim = c(0.9,1.1))
```
